Cheminformatics: The Computer Science of Chemical Discovery

### 2011-11-12 rebuttal.tex

We would like to thank the anonymous reviewers for their constructive
comments. We have updated the main text in several places and have
also included a supplemental document that provides a brief
description of terminology and a list of some historic milestones in
the field. The following sections provide a detailed response to
individual reviewer comments.
\vspace{2em}

\fbox{\textbf{Reviewer 1}}

\comment{when the authors state 'together with an even larger number
  of annotations', they shold explain what are annotations and how
  much is the large number;}
\response{We have updated the text on page 1 to be more explicit in
  terms of what annotations are and provide a concete example of the
  sizes involved.}

\comment{For an overview I would expect a more comprehensive set of
  references.} 
 \response{We agree with the reviewer that a more
  comprehensive list of references would be appropriate and
  useful. However, due to format limitations of the article, we have
  been restricted to the use of 40 references. To alleviate this
  problem, we have included a supplement that provides a more
  extensive list of references, specifically including a number of
  texts covering cheminformatics broadly.}

\comment{1) The connection between sections could be improved,
  i.e. the authors should try to enhance the correlation that exists
  between the topics described in each section and map them to a
  global flow of cheminfomatics changes. 2) The authors state in the
  beginning that the paper will be around the concept of risk
  minimization in drug discovery, but I found that this topic is not
  sufficiently explored. In many sections is not straightforward teh
  connection between the challenges described and their impact on risk
  minimization in drug discovery; 3) The paper misses some important
  efforts and more detail on translational medicine and semantic web }
\response{The reviewer is correct in noting the disjointed character
  of the original article. We have updated the text throughout to flow
  smoothly. As part of this we have made more explicit how each topic
  that is discussed plays a role in derisking various stages of the
  drug disocvery process. Finally, we have updated the text on page 4
  to address the issue of linked data and the role that semantic
  technologies can play in this area. We have, however, chosen to not
  address the issue of translational medicine. While there have been
  some cases where cheminformatics have played a role in translational
  medicine, the bulk of cheminformatics lies upstream of this
  topic. Hence, to keep the article focused we have chosen to not
  include this topic.}

### 2011-11-12 paper.tex
Updated text to complete response to reviewer 1

The preceding section focuses on the choice of representation of
chemical structures. With the profusion of databases that contain
details of small molecules, an emerging problem is the integration of
such databases. Such integration forms the basis of ``systems'' level
investigations of biological systems. A current challege in
cheminformatics is how to identify ``equivalent'' objects across data
sources. At one level, equivalent compounds can be identified via
canonical structure representations (InChI). But there are other
equivalences beyond structure. For example, compounds may bind to the
same or similar targets; they may be have similar biological function
(say, anti-neoplastic) and so on. To assess such higher level
characteristics, ontologies and semantic technologies have been
employed, to first, support annotations of chemical systems and
second, enable semantic analyses of chemical systems in conjunction
with other biological systems (such as pathways). A number of
challenges remain in this area - the use of ontologies allows one to
build a semntic network of concepts that links molecules to other
molecules as well as other objects such as targets, genes and
pathways. How does one evaluate similarities in such a heterogenous
network \cite{couto2010}? How does one integrate chemical structure
information with ontologies \cite{hastingsowled2010}? Integrating
small molecule data with other biological data types will be key to
acheiving a systems level understanding of small
molecules \cite{Oprea:2007fk}.

>>>>

The preceding section focuses on the choice of representation of
chemical structures. With the profusion of databases that contain
details of small molecules, an emerging problem is the integration of
such databases. Such integration forms the basis of ``systems'' level
investigations of biological systems. A current challege in
cheminformatics is how to identify ``equivalent'' objects across data
sources. At one level, equivalent compounds can be identified via
canonical structure representations (InChI). But there are other
equivalences beyond structure. For example, compounds may bind to the
same or similar targets; they may be have similar biological function
(say, anti-neoplastic) and so on. To assess such higher level
characteristics, ontologies and semantic technologies have been
employed, to first, support annotations of chemical systems and
second, enable semantic analyses of chemical systems in conjunction
with other biological systems (such as pathways). A number of
challenges remain in this area - the use of ontologies allows one to
build a semntic network of concepts that links molecules to other
molecules as well as other objects such as targets, genes and
pathways. How does one evaluate similarities in such a heterogenous
network \cite{couto2010}? How does one integrate chemical structure
information with ontologies \cite{hastingsowled2010}? And more
generally, how can we achieve automated inferencing in such
heterogenous linked systems, allowing us to acheive a systems level
understanding of small molecules \cite{Oprea:2007fk}.
